# R Bootcamp Tutorial

## Debugging

Each of the following code blocks contain one or more errors. You should:

1. Attempt to run the code
2. Diagnose the problem
3. Correct the problem  
4. Add a bit of documentation to the code
5. Based on what you now know about R, try to create your own debugging exercise

Some notes:

- Remember to restart your R session before attempting these exercises (`Ctrl+Shift+F10`)
- Pay attention to any error messages; they are trying to help you!
- Remember you have the internet at your disposal
- Not every problem will produce an error message

> Thank you to Thomas Wood and the R Python User Group for providing many debugging examples!

1. 

```{r}
list.files("C:\Program Files (x86)")
```

2. 

```{r}
аbs(-1)
```

3. 

```{r}
x <- "hello"
y <- "goodbye"
z <- x + y

print(z)
```

4. 

```{r}
select(.data = mtcars, cyl)
```

5. 

```{r}
library(dplyr)
num_cyl <- 6.0
filter(.data = mtcars, num_cyl == num_cyl)
```


6. 

```{r}
library(purrr)
map(.x = 1:2, .f = rnorm(n = 10, mean = .x))
```
