# R Bootcamp Exercise

## Analyzing Data

Remember to load in your packages!

```{r, include = FALSE}

# Load libraries
library(cansim) # read in CODR/NDM tables
library(dplyr)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

## Load your data

1. Load your cleaned data from [the last exercise](./2_cleaning_data.Rmd) or run this code chunk.

```{r}

# Load our data 
df <- get_cansim("32-10-0112-01")

#drop totally empty columns
for (col in colnames(df)){      # loop through all our columns
    if (all(is.na(df[col]))){  # check if NAs appear in the respective column in ALL rows
      df <- select(df, -(col))  # drop that column from our dataframe
      }              
}
#drop boring columns

df <- select(df, -(ends_with("ID") | "STATUS" | "DECIMALS" | "TERMINATED"))

```

## Analysis

2. Does any commodity have a value of exactly 999 units at any point in time? Which, when and where?

```{r}


```

3. Filter for your favourite dairy product. Compute the mean, standard deviation, min, max, and any other relevant summary statistics for this product.

```{r}


```

4. Find out what is the most numerous commodity in Canada (or another region) over the years.

```{r}


```

5. Compute the cheese stock per capita over time:
Access population data [here](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710000501).

```{r}


```
