# R Bootcamp Exercise

## Visualizing Data

Remember to load in your packages!

```{r, include = FALSE}

# Load libraries
library(cansim) # read in CODR/NDM tables

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

## Load your data

1. Load your cleaned data from [the last exercise](./2_cleaning_data.Rmd) or run this code chunk.

```{r}

# Load our data 
df <- get_cansim("32-10-0112-01")

#drop totally empty columns
for (col in colnames(df)){      # loop through all our columns
    if (all(is.na(df[col]))){  # check if NAs appear in the respective column in ALL rows
      df <- select(df, -(col))  # drop that column from our dataframe
      }              
}
#drop boring columns

df <- select(df, -(ends_with("ID") | "STATUS" | "DECIMALS" | "TERMINATED"))

```

## Visualizations

For this exercise, create a visualization of your choice (or use the suggestion) for each question from [3_analyzing_data.Rmd](./3_analyzing_data.Rmd)

2. Does any commodity have a value of exactly 999 units at any point in time? Which, when and where?

Suggestion: Create a pie chart showing which Commodities or Regions fit the above criteria.

```{r}


```

3. Filter for your favourite dairy product. Compute the mean, standard deviation, min, max, and any other relevant summary statistics for this product.

Suggestion: Plot the mean as a bar graph over time. For a bonus, plot error bars as well. See [here](https://r-graph-gallery.com/4-barplot-with-error-bar.html) for a guide.

```{r}


```

4. Find out what is the most numerous commodity in Canada (or another region) over the years.




Suggestion: Create an animated bar chart showing the values of the top 3 commodities over time.

> Hint: you may need to adjust your code to get the top 3 commodities, rather than the top 1, for each group!

```{r}


```

5. Compute the cheese stock per capita over time:
Access population data here:https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710000501

Suggestion 1. Plot a line graph of the percapita cheese stock over time.
Suggestion 2. Create a plot with two lines and two axes: one for the cheese stock and the second for the population over time.

```{r}


```
