# R Bootcamp Tutorial

## Debugging

Each of the following code blocks contain one or more errors. You should:

1. Attempt to run the code
2. Diagnose the problem
3. Correct the problem  
4. Add a bit of documentation to the code
5. Based on what you now know about R, try to create your own debugging exercise

Some notes:

- Remember to restart your R session before attempting these exercises (`Ctrl+Shift+F10`)
- Pay attention to any error messages; they are trying to help you!
- Remember you have the internet at your disposal
- Not every problem will produce an error message

> Thank you to Thomas Wood and the R Python User Group for providing many debugging examples!

1. 

```{r}
list.files("C:\Program Files (x86)")
```

```{r}
# This line was missing a second \
list.files("C:\\Program Files (x86)")
```

2. 
```{r}
library(purrr)
map(.x = 1:2, .f = rnorm(n = 10, mean = .x))
```

```{r}
#This chunk was improperly using purrr techniques. The syntax was incorrect.
1:2 %>%
  map(rnorm, n=10)

#To maintain the style of the original, you must use an anonymous function
1:2 %>%
  map(function(x) rnorm(10, x))

#third option
map(.x = 1:2, .f = function(.x){rnorm(n = 10, mean = .x)})
```

3. 

```{r}
аbs(-1)
```

```{r}
#this line had an improperly encoded symbol
abs(-1)
```

4. 

```{r}
x <- "hello"
y <- "goodbye"
z <- x + y

print(z)
```

```{r}
# this chunk tried to use numerical methods for characters. There are multiple solutions depending on the desired outcome
x <- "hello"
y <- "goodbye"
z <- c(x,y)

print(z)
```

5. 

```{r}
select(.data = mtcars, cyl)
```

```{r}
#This chunk did not have the required package loaded
library(dplyr)
select(.data = mtcars, cyl)
```

6. 

```{r}
library(dplyr)
num_cyl <- 6.0
filter(mtcars, num_cyl == num_cyl)
```

```{r}
#The column name in the filter was not correct
num_cyl <- 6
filter(.data = mtcars, cyl == num_cyl)
```
