# R Bootcamp Tutorial Exercises

Remember to load in your packages!

```{r, include = FALSE}
# Load libraries
library(cansim) # read in CODR/NDM tables

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Load your data

1. Load in yet another dairy dataset from here [this data set](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3210011201). Alternatively, you may browse the [Government of Canada Open Data Portal](https://www150.statcan.gc.ca/n1/en/type/data?MM=1#tables) for a dataset on a topic that interests you.

```{r}
# Load the cansim package
library(cansim)

# Load our data 
df <- get_cansim("32-10-0112-01")
```

## Get to know your data

1. What are the columns?

```{r}
colnames(df)
summary(df)
```

2. What are the data types?

```{r}
summary(df)
sapply(df, typeof)
sapply(df, class)
```

3. What are some values in your columns?

```{r}
summary(df)

df_cols <- colnames(df)
for (col in df_cols){
   print(unique(df[col]))
}
```

1. What is most interesting to you about your data?
