# R Bootcamp Tutorial

## Programming Logic (Optional)

In this lesson, we will cover:

1. Objects and Variables
2. Data types
3. Loops
4. Vectorization
5. Functions

## 1. Objects and Variables

R is an object-oriented programming language. That means that it allows you to store single values (think a number or text) as a "variable." This is analogous to putting an object in a box and labeling that box. As opposed to SAS, which is a (premdominantly) procedural programming language.

You can use a caret and dash to assign an object a value. For example:

```{r}
surveys <- 20 
```

`print` will allow you to see what is stored in a variable.

```{r}
print(surveys)
```

## 2. Data Types

There are different data types in programming languages. Different types of data are treated differently by R and its functions. In R, the simplest are:

* Character: Comprised of text with quotations around it. Ex: "red"
* Numeric: Comprised of numbers
* Logical: TRUE or FALSE

Try the following:

```{r}
class(surveys)
```

What type of variable is `surveys`?

## 3. Loops

One great thing about computers is that they don't mind completing repetitive tasks that might not be so fun for a human, leading to manual errors and boredom. *Loops* are iterative processes that apply instructions over a predefined set of data.

*For loops* are some of the most common and useful types of loops in R. Take a look at this code:

```{r}
for (i in 1:5) {
  print(i)
}
```

The loop defines a vector (1:5), which contains all of the numbers from 1 to 5. On each iteration of the loop, `i` represents one of these numbers. Then, the loop is instructed to print each `i`. Run the code and see what it produces! What would changing the vector's length do?

You might want to use a loop to perform some operation rather than just printing. The method is the same:

```{r}
for (i in 1:5) {
  i <- i+2
  print(i)
}
```

## 4. Vectorization

Vectorization is an alternative to loops. It is typically more efficient and can greatly condense your code.

> Whenever possible you should to try avoid using loops.

We can rewrite the previous code block as:

```{r}
vec <- 1:5

vec <- vec+2
print(vec)
```

The most common methods in R are nearly all vectorized (or vectorizable with a bit of work!)

## 5. Functions

Functions allow you to write a generic bit of code that can be reused throughout your script (or even pulled into other scripts!)

Let's turn our previous code block into a function.

```{r}
add_two <- function(x){
  return(x+2)
}

print(add_two(1))
print(add_two(1:5))
print(add_two(vec))
```
